---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example engine (uses the RStudio API to call the terminal)

Firstly, `cling` returns the errors colored with ANSI shell color codes. To return the colors, instead of the codes, we follow a solution posted in [crayon#24](https://github.com/r-lib/crayon/issues/24#issuecomment-581068792)
.
```{r color, echo = TRUE, results='asis'}
# https://github.com/r-lib/crayon/issues/24#issuecomment-581068792
# crayon needs to be explicitly activated in Rmd
options(crayon.enabled = TRUE)
# Hooks needs to be set to deal with outputs
# thanks to fansi logic
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
                                   which = c("output", "message", "error"))
```

Now, we can define our language engine:

```{r}
# Note: RStudio needs to be running ...
# https://rstudio.github.io/rstudio-extensions/rstudioapi.html
# https://github.com/rstudio/rstudio/issues/6892
#SOLUTION: Run in RStudio CONSOLE: rmarkdown::render("my_rmd_code_file.Rmd")
# Create a global variable for the terminal:
while(length(rstudioapi::terminalList()) > 0){
  rstudioapi::terminalKill(rstudioapi::terminalList()[1])
}
ClingRStudioTerminal <<- rstudioapi::terminalCreate(caption = "RStudioClingTerminal", 
                                                    show = FALSE)
rstudioapi::terminalSend(ClingRStudioTerminal, "cling\n") 

# The knitr_engine for cling:
knitr::knit_engines$set(cling = function(options) {
  if(options$eval){
    # If we want to clear the environment - close the terminal and create a new one:
    if(sum(options$engine.opts$cling == 'ClearClingEnv') == 1){
      # Quit cling:
      rstudioapi::terminalSend(ClingRStudioTerminal, ".q\n")
      # Close the terminal
      rstudioapi::terminalKill(ClingRStudioTerminal)
      # Create a new terminal instance
      ClingRStudioTerminal <<- rstudioapi::terminalCreate(caption = "RStudioClingTerminal", 
                                                      show = FALSE)
      # wait for it to start
      while (!rstudioapi::terminalRunning(ClingRStudioTerminal)) {
        Sys.sleep(0.1)
      }
      # Start cling
      rstudioapi::terminalSend(ClingRStudioTerminal, "cling\n")
    }
    # https://cran.r-project.org/web/packages/rstudioapi/vignettes/terminal.html
    if (!rstudioapi::terminalRunning(ClingRStudioTerminal)) {
      # start the terminal shell back up, but don't bring to front
      rstudioapi::terminalActivate(ClingRStudioTerminal, show = FALSE)
      # wait for it to start
      while (!rstudioapi::terminalRunning(ClingRStudioTerminal)) {
        Sys.sleep(0.1)
      }
    }
    # Clear the current terminal window:
    rstudioapi::terminalClear(ClingRStudioTerminal)
    
    # https://stackoverflow.com/a/25105442
    # Redirect cling output to a temporary txt file:
    #rstudioapi::terminalSend(ClingRStudioTerminal, ".&> RStudioClingOutputTMP.txt\n")
    rstudioapi::terminalSend(ClingRStudioTerminal, ".1> RStudioClingOutputTMP_stdout.txt\n")
    rstudioapi::terminalSend(ClingRStudioTerminal, ".2> RStudioClingOutputTMP_sterr.txt\n")
    code <- paste(options$code, collapse = '\n')
    rstudioapi::terminalSend(ClingRStudioTerminal, paste0(code, "\n"))
    
    # To obtain the results programmatically, wait for cling to finish saving:
    while(tail(rstudioapi::terminalBuffer(ClingRStudioTerminal), 1) != "[cling]$ "){
      Sys.sleep(0.1)
    }
    
    suppressWarnings({
      # Read the output:
      out <- readLines("RStudioClingOutputTMP_stdout.txt")
      out <- paste(out, collapse = '\n')
      # Check if there are any errors:
      err <- readLines("RStudioClingOutputTMP_sterr.txt") 
      err <- paste(err, collapse = '\n')
    })
    if(length(err) > 0 & err != ""){
      if(!options$error) stop(err)
      out <- paste0(out, ifelse(out != "", "\n", ""),
                    paste0(rep("~", 80), collapse = ""),
                    "\n\nknitr: Errors while running chunk:\n", 
                    err,
                    paste0(rep("~", 80), collapse = ""),
                    "\n\nknitr: new variables NOT saved in session, however, existing variable values changed!")
      # Use .undo [n] to UNDO the input lines from this chunk:
      rstudioapi::terminalSend(ClingRStudioTerminal, paste0(".undo ", length(options$code), "\n"))
      # TODO: maybe undo only if !options$error ?
    }
    # Remove the file
    # writeLines("", "RStudioClingOutputTMP.txt")
  }else{
    out <- ""
  }
  
  # Return the results:
  knitr::engine_output(options, options$code, out)
})
```

## Example call to the engine:

Call the engine a couple of times with code clearing:

```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```

Next, call one without any code clearing:

```{cling}
printf("Testing...\n");
```

```{cling}
printf("Testing...\n");
```

Next, clear the environment again (se the `engine.opts` in the `.Rmd` file for the code chunk below), and run the code below:

```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling}
int x;
```

```{cling}
x
```


```{cling}
x += 1;
x
```

## Example - with error

```{cling, engine.opts = list(cling = 'ClearClingEnv'), error = TRUE}
int x = 1;
x
int x = 2;
x
y
```

(Note: it is technically possible to allow variable redefinition, see the [The LLVM Project Blog: "Interactive C++ with Cling"](https://blog.llvm.org/posts/2020-11-30-interactive-cpp-with-cling/)).

Test to make sure that this chunk was undone - variable `x` should not exist in current session:

```{cling, error = TRUE}
x
```
Create a new variable (no errors - this should persist)

```{cling, error = TRUE}
int y = 1;
y
```

Another chunk - new value for `y`, define a new variable `z` and an error for an undefined `x`:

```{cling, error = TRUE}
y += 1;
int z = 2;
x
```

As a result, the previous chunk is undone, `z` does not exist, however `y` value remains changed, since it was created in a previous chunk!

```{cling, error = TRUE}
y
```
```{cling, error = TRUE}
z
```

See [ROOT files](https://root.cern.ch/root/html534/guides/users-guide/InputOutput.html) for an idea on the possible solutions for variable cache creation.

**Important!** knit this file by running `rmarkdown::render("00_cling_language_engine_knitr.Rmd")` inside the RStudio console.
