---
title: "Example: calling `cling`"
output: 
  html_document:
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Important!** knit this file by running `rmarkdown::render("00_cling_language_engine_knitr.Rmd")` inside the RStudio console.

## Example engine (uses the RStudio API to call the terminal)

Firstly, `cling` returns the errors colored with ANSI shell color codes. To return the colors, instead of the codes, we follow a solution posted in [crayon#24](https://github.com/r-lib/crayon/issues/24#issuecomment-581068792)
.
```{r color, echo = TRUE, results='asis'}
# https://github.com/r-lib/crayon/issues/24#issuecomment-581068792
# crayon needs to be explicitly activated in Rmd
options(crayon.enabled = TRUE)
# Hooks needs to be set to deal with outputs
# thanks to fansi logic
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
                                   which = c("output", "message", "error"))
```

Next, we load our language engine:

```{r}
source("../src/cling.R")
```

## Example call to the engine:

Call the engine a couple of times with code clearing:

```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```

Next, call one without any code clearing:

```{cling}
printf("Testing...\n");
```

```{cling}
printf("Testing...\n");
```

Next, clear the environment again (se the `engine.opts` in the `.Rmd` file for the code chunk below), and run the code below:

```{cling, engine.opts = list(cling = 'ClearClingEnv')}
#include <stdio.h>

printf("Hello World!\n");
```


```{cling}
int x;
```

```{cling}
x
```


```{cling}
x += 1;
x
```

## Example - with error

If an error is encountered - the lines in the code chunk are undone with the `.undo [n]` command. Below contains an example, when we try to define `x` twice, as well as call an undefined variable `y`

```{cling, engine.opts = list(cling = 'ClearClingEnv'), error = TRUE}
int x = 1;
x
int x = 2;
x
y
```

(Note: it is technically possible to allow variable redefinition, see the [The LLVM Project Blog: "Interactive C++ with Cling"](https://blog.llvm.org/posts/2020-11-30-interactive-cpp-with-cling/)).

Test to make sure that this chunk was undone - variable `x` should not exist in current session:

```{cling, error = TRUE}
x
```
Create a new variable (no errors - we expect this variable to persist if there is an error in later chunks):

```{cling, error = TRUE}
int y = 1;
y
```

In a new chunk - assign a new value for `y`, define a new variable `z` and try to call an undefined `variable x`:

```{cling, error = TRUE}
y += 1;
int z = 2;
x
```

As a result, the previous chunk is undone, `z` does not exist, **however `y` value remains changed, since it was created in a previous chunk!**

```{cling, error = TRUE}
y
```
```{cling, error = TRUE}
z
```

See [ROOT files](https://root.cern.ch/root/html534/guides/users-guide/InputOutput.html) for an idea on the possible solutions for variable cache creation.