---
title: "Example: `cling` with `libgsl`"
output: 
  html_document:
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Important!** knit this file by running `rmarkdown::render("01_libgsl.Rmd")` inside the RStudio console.

## Example engine (uses the RStudio API to call the terminal)

Firstly, `cling` returns the errors colored with ANSI shell color codes. To return the colors, instead of the codes, we follow a solution posted in [crayon#24](https://github.com/r-lib/crayon/issues/24#issuecomment-581068792)
.
```{r color, echo = TRUE, results='asis'}
# https://github.com/r-lib/crayon/issues/24#issuecomment-581068792
# crayon needs to be explicitly activated in Rmd
options(crayon.enabled = TRUE)
# Hooks needs to be set to deal with outputs
# thanks to fansi logic
old_hooks <- fansi::set_knit_hooks(knitr::knit_hooks, 
                                   which = c("output", "message", "error"))
```

Load our language engine:

```{r}
source("../src/cling.R")
```

## Installing `libgsl`

Easiest way to install the library, as described [here](https://askubuntu.com/a/1052165):

```{bash, eval = FALSE}
sudo apt-get install libgsl-dev
```

## Using the third-party libraries in `cling`

We need to tell the interpreter where the third-party libraries are located. Following [xeus-cling#87](https://github.com/jupyter-xeus/xeus-cling/issues/87#issuecomment-349053121), the are three ways to include either specific libraries, oradding the path as well:

- If you want to add a path to the include path:
```{cc eval = FALSE}
#pragma cling add_include_path("inc_directory")
```
- If you want to add a path to the library path:
```{cc, eval = FALSE}
#pragma cling add_library_path("lib_directory")
```
- If you want to load a library:
```{cc, eval = FALSE}
#pragma cling load("libname")
```
Note that you do not need to add the library extension (`.so`, '.lib`, or .`dylib`).

A similar answer is given [here](https://stackoverflow.com/a/50291733). Basically, the binary code is loaded with `#pragma cling load(...)`, and `#include`

## Example 1: using `libgsl` and `cling`

Load the required libraries:

```{cling, engine.opts = list(cling = 'ClearClingEnv'), error = TRUE}
/* Tell cling to link the libraries */
#pragma cling load("gsl")
#pragma cling load("gslcblas")

/* Load the libraries */
#include <stdio.h>
#include <gsl/gsl_statistics.h>
```

Create a data sample:

```{cling}
double data[5] = {17.2, 18.1, 16.5, 18.3, 12.6};
```

```{cling}
printf("The dataset is %g, %g, %g, %g, %g\n",
       data[0], data[1], data[2], data[3], data[4]);
```

Calculate some statistics for this data sample:

```{cling}
double mean, variance, largest, smallest;

mean     = gsl_stats_mean(data, 1, 5);
variance = gsl_stats_variance(data, 1, 5);
largest  = gsl_stats_max(data, 1, 5);
smallest = gsl_stats_min(data, 1, 5);
```

Print the results:

```{cling}
printf("The sample mean is %g\n", mean);
printf("The estimated variance is %g\n", variance);
printf("The largest value is %g\n", largest);
printf("The smallest value is %g\n", smallest);
```

# Example 2 (not recommended)

A `cling`-specific [metaprocessor](https://root.cern/cling/#metaprocessor) flag allows loading a library interactively (as described [here](https://stackoverflow.com/a/45679612)):

```{cling, engine.opts = list(cling = 'ClearClingEnv'), error = TRUE}
.L gsl

#include <stdio.h>
#include <gsl/gsl_statistics.h>
```

```{cling}
double data[3] = {1.0, 2.0, 3.0};
printf("The sample mean is %g\n", gsl_stats_mean(data, 1, 3));
```

The difference is that `Example 1` uses the [pragma](https://docs.microsoft.com/en-us/cpp/preprocessor/pragma-directives-and-the-pragma-keyword?view=msvc-160) directive, which is a compiler specific feature and is a bit closer to most general `C++` examples.

# Example 3 (external file, experimental, not run)

This example is more of a general idea, which follows [this answer on ROOT](https://root-forum.cern.ch/t/loading-libraries-in-a-cling-script/19973/2)

The idea is to pack all of the preliminary libraries into a "script" of sorts. For example:

- Create a file `prepareWorkspace.C` with the following code:

```{c, eval = FALSE}
#include "cling/Interpreter/Interpreter.h"

void prepareWorkspace() {
  gCling->loadFile("libgsl");
  /*
    Add any other libraries here, if needed.
  */
}
```
This file should be in the same directory, as the terminal, opened by `rstudioapi`.

- Then, similarly to how we used `.L`, we now use the `.x` command:

```{cling, eval = FALSE}
.x prepareWorkspace.C
```

From the [documentation](https://root.cern/cling/#metaprocessor):

> `.x filename.cxx` - loads filename and calls `void filename()` if defined